---
title: "Assessment 1.1"
author: "CAO YUAN 4659774"
date: "8/25/2021"
output: html_document
---


install.packages("tidyverse")  
library(tidyverse)
install.packages("knitr")  
library(knitr)
install.packages("skimr")  
library(skimr)
install.packages("dplyr")
library(dplyr)
install.packages("readr")
library(readr)

# ***RMarkdown editing***
# ***R Markdown provides an authoring framework for data science. You can use a single R Markdown file to save and execute code and generate high quality reports that can be shared with an audience.R Markdown documents are fully reproducible and support dozens of static and dynamic output formats***

[links](https://rmarkdown.rstudio.com/lesson-8.html)


```{r}
fastfood<-read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
```



head(fastfood, n=10)

kable(fastfood[1:10, 1:5])


  
```{r}
fastfood %>%
  filter(calories > 1000) %>%
   knitr::kable()
```

```{r}
  fastfood %>%
  filter(total_fat > 40 & total_carb > 80) %>%
  as_data_frame()->dont_eat_this
  kable()
```
   

```{r}
fastfood %>%
filter(restaurant == 'Mcdonalds') %>%
   knitr::kable()
```

### Part 2

library(skimr)
skim(fastfood)


```{r}
fastfood %>%
  count(restaurant) %>%
  kable()
```

```{r}
fastfood %>%
  group_by(restaurant) %>%
  summarise(average_cal = mean(calories)) %>%
  kable()
```


```{r}
 fastfood %>%
  group_by(restaurant) %>%
  mutate(average_calories = mean(calories)) %>%
  ungroup() %>%
 kable()
  
```
```{r}
fastfood %>%
  ggplot(., aes(restaurant, total_fat)) +
  geom_col()
```


```{r}
 fastfood %>%
  group_by(restaurant) %>%
  mutate(max_total_fat = max(total_fat)) %>%
  ungroup() %>%
 kable()
  
```
install.packages("ggplot2")
library(ggplot2)

```{r}
 fastfood %>%
  filter(restaurant == 'Mcdonalds') %>%
  ggplot(., aes(sugar, protein)) +
   geom_point()
```
### Part 3
```{r}
fastfood$restaurant<-as_factor(fastfood$restaurant) 
class(fastfood$restaurant)
```

calories_type<-ordered(cut_number(fastfood$calories, 3, labels = c("low", "med","high")))
fastfood<-cbind(fastfood,calories_type)

library(ggplot2)

ggplot(fastfood, aes(restaurant, fill=factor(calories_type))) + geom_bar(position='stack')

trans_fat_percent<-round(fastfood$trans_fat/fastfood$total_fat*100,2)
fastfood<-cbind(fastfood,trans_fat_percent)


ggplot(fastfood, aes(restaurant, trans_fat,  color = restaurant, fill = restaurant)) + geom_point(alpha=0.4, position = "jitter")

```{r}
fastfood %>%
  group_by(restaurant) %>%
  summarise(average_fat = mean(total_fat)) %>%
  kable()
```

ggplot(fastfood, aes(restaurant, total_fat, fill = restaurant)) + geom_boxplot()
