---
title: "Assignment2.rmd"
author: "CAO YUAN 4659774"
date: "9/2/2021"
output: html_document
---

install.packages("tidyverse")  
library(tidyverse)
install.packages("knitr")  
library(knitr)
install.packages("skimr")  
library(skimr)
install.packages("dplyr")
library(dplyr)
install.packages("readr")
library(readr)
install.packages("tidymodels")
library(tidymodels)
install.packages("plyr")
library(plyr)
library(ggplot2)
library(patchwork)
library(forcats)
library(tidyr)
install.packages("imager")
```{r}
library(tidyverse)
library(dplyr)
library(skimr)
library(knitr)

```


```{r}
thanksgiving_meals<-readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-20/thanksgiving_meals.csv')

```
img<-load.image (Users/rubycao/Desktop/ruby.JPG)
###Part 2: Data Wrangling and visualization

### 1
```{r}
 thanksgiving_meals %>%
  head(10) %>%
 kable()
```
### 2
```{r}
thanksgiving_meals %>%
  mutate(family_income=fct_reorder(family_income, parse_number(family_income))) %>%

  kable()
```
### 3
```{r}
skimr::skim(thanksgiving_meals) %>%
  kable()
```

```{r}
summary(thanksgiving_meals) %>%
  kable()
```

### 4
```{r}
thanksgiving_meals %>%
  filter(celebrate == 'Yes') %>%
   count()
```
### 5
```{r}
thanksgiving_meals %>% filter(celebrate != 'No') %>%
  dplyr::count(main_dish, main_prep,sort=TRUE) %>%
  kable()
```
### 6
```{r}
thanksgiving_meals %>% filter(celebrate != 'No') %>% count(main_dish, main_prep, sort=TRUE) %>% 
ggplot(aes(n,main_dish,fill=factor(main_prep))) +  geom_col(position='stack',alpha=0.8) +coord_flip() +theme_linedraw()
```
### 7
```{r}
thanksgiving_meals %>% filter(cranberry!='None') %>%
   count() %>%
  kable()
```

```{r}
thanksgiving_meals %>% filter(gravy == 'Yes') %>%
   count() %>%
  kable()
```
### 8/9
```{r}
thanksgiving_meals %>%
  mutate(family_income=fct_reorder(family_income, parse_number(family_income))) %>%
  dplyr::count(celebrate,family_income,sort=TRUE) %>%
  dplyr::filter(celebrate == 'Yes') %>%
  group_by(family_income) %>%
  ggplot(aes(family_income,n,fill=celebrate)) +
  geom_col() +
  coord_flip() +
  labs(x="Income range", y="Number of people") +
 theme_test()
```

###10/11
```{r}
thanksgiving_meals %>%
select(starts_with("side"),-side15) %>%
  gather(type, value,na.rm=TRUE) %>%
   dplyr::count(type,value, sort = TRUE) %>%
  select(value, n) %>%
  head(1) %>%
  as_data_frame()->most_common_side 
  kable()
```

```{r}
thanksgiving_meals %>%
select(starts_with("pie"),-pie13) %>%
  gather(type, value,na.rm=TRUE) %>%
   dplyr::count(type,value, sort = TRUE) %>%
  select(value, n) %>%
  head(1) %>%
  as_data_frame(most_common_pie)%>%
  kable()
```

```{r}
thanksgiving_meals %>%
select(starts_with("dessert"),-dessert11) %>%
  gather(type, value,na.rm=TRUE) %>%
   dplyr::count(type,value, sort = TRUE) %>%
  filter(value!="None") %>%
  select(value, n) %>%
  head(1) %>%
  as_data_frame(most_common_dessert)%>%
  kable()
```

```{r}
 union(most_common_dessert, most_common_pie)%>% union(most_common_side) %>%
  ggplot(aes(value,n,fill=value)) +
  geom_col() +
   labs(title= "Most common dessert, side and pie") +
  theme_test()
  
```


###12/13
```{r}
thanksgiving_meals %>%
  mutate(family_income=fct_reorder(family_income, parse_number(family_income))) %>%
  dplyr::filter(celebrate == 'Yes') %>%
  select(family_income, main_dish) %>%
   dplyr::count(family_income, main_dish) %>%
  kable()
```

```{r}
thanksgiving_meals %>%
  mutate(family_income=fct_reorder(family_income, parse_number(family_income))) %>%
  dplyr::filter(celebrate == 'Yes') %>%
  select(family_income, main_dish) %>%
   dplyr::count(family_income, main_dish) %>%
  ggplot(aes(family_income,n,fill=main_dish)) +
  geom_col() +
  coord_flip() +
  labs(x="Income range", y="Number of people") +
 theme_test()

```


